MSDS 7333 Case Study 06

### 2017-06-21 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Added figure and others. See desc.
 I added a KNN figure, fixed the tables, removed comments, corrected some spelling/grammar, and added a "Future Study" section which describes some of what Matt was talking about

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Table of Contents: {#TOC}
<li>[Assignment Requirements](#AssignmentOutline)</li>
<li>[Introduction](#Introduction)</li>
<li>[**Part 1**](#Part1)</li>
<ul> 
  <li>[Signal Collection](#SigCol)</li>
  <li>[Signal Strength](#SigStren)</li>
  <li>[Signal Density](#SigDen)</li>
  <li>[Discussion](#Discussion)</li>
  <li>[k-Nearest Neighbors](#knn)</li>
</ul>
<li>[**Part 2**](#Part2)</li>
<ul> 
  <li>[2.1 Examining Power](#ExamPower)</li>
  <li>[2.2 Examining Weights](#ExamWeight)</li>
  <li>[2.3 Weighted kNN](#WtKNN)</li>
</ul>
<li>[References](#References)</li>
<li>[Appendix](#Appendix)</li>

@@

<font color = "purple"><b>>>>>Everything in this color to be removed at final review<<<<</b></font>

<font color = "purple"><b>
- Discuss the goal of the case study
</b></font>

@@

<font color = "red"><b>
- Need example training input line

t=1139643118358;id=00:02:2D:21:0F:33;pos=0.0,0.0,0.0;degree=0.0;00:14:bf:b1:97:8a=-38,2437000000,3;00:14:bf:b1:97:90=-56,2427000000,3;00:0f:a3:39:e1:c0=-53,2462000000,3;00:14:bf:b1:97:8d=-65,2442000000,3;00:14:bf:b1:97:81=-65,2422000000,3;00:14:bf:3b:c7:c6=-66,2432000000,3;00:0f:a3:39:dd:cd=-75,2412000000,3;00:0f:a3:39:e0:4b=-78,2462000000,3;00:0f:a3:39:e2:10=-87,2437000000,3;02:64:fb:68:52:e6=-88,2447000000,1;02:00:42:55:31:00=-84,2457000000,1

- Need table of training data: name, description, type, example

|Data Element|Description|Type|Example|
|:-|:-|:-|:-|
|t |timestamp (milliseconds since midnight 1/1/1970)|long integer|t=1139643118358 --> 2/11/2006 07:35:28|
|id|target device MAC address|6-octet sequence in hexadecimal|id=00:02:2D:21:0F:33|
|pos|coordinates of the target device in meters from a choke point in the host building|3 floating point values, comma-separated|pos=0.0,0.0,0.0|
|degree|Orientation of the target device (with respect to North) in degrees|float|degree=0.0|
|MAC Address|one or more reports from various Access Points|comma-separated value list|00:14:bf:b1:97:8a=-38,2437000000,3|

|Data Element|Description|Type|Example|
|-|-|-|-|
|signal|The signal strength reported for the Access Point in dBm|integer|-38|
|channel|The frequency of the channel used by the Access Point in Hz|integer|2437000000|
|mode|The mode of operation for the Access Point (1 = Ad Hoc, 3 = Access Point)|integer|3|

</b></font>

<font color = "purple"><b>
- Discuss the different Access points and MAC Addresses
</b></font>

Six APs were initially discovered in the original analysis in the case study. The authors then decided to reject AP "00:0f:a3:39:dd:cd" (rejectedAP) stating it was identified as an extra address. The details of their reasoning was not clearly explained. The authors chose AP "00:0f:a3:39:e1:c0" (chosenAP) instead. This case study will evalute the author's decision and decide if the IPS system would be more accurate with the chosenAP, rejectedAP, or both APs.

<font color = "purple"><b>
- BRIEF overview of K-Nearest Neighbors
</b></font>

@@

<font color = "red"><b>
- can we still a graphic from scikit-learn of k-NN model?
</b></font>

In part 1 of the case study, the author's original decision to reject an AP is revisited to see if this was the best decision. In part 2, the prediction of X,Y based on the average of its neighbors is revisited and modified to a weighted sum of X,Y to determine if this change will produce better predictions. Alternate weights will also be analyzed.

@@

```{r}
plotSignalMaps(c(chosenAP, rejectedAP))
#  figure 1 - number of reports at each location for the rejected and chosen APs.
```

@@

```{r}
df = subset(offline, posX ==2 & posY == 12 & !(mac %in% remainingAP), c(signal, angle, mac))
plotBoxplotSignalStrength(df)
# figure 2 - Signal strength distribution across both Access Points
```

The higher the decibel-milliwatts means a stronger signal strength which indicates a closer physical position to the AP since signal strength decreases exponentially with distance.<font color="red">Find this in the book and cite it.</font> The distribution shows AP "e1:c0" has a stronger signal strength at this location than "dd:cd." But, the two of them together cover a wider range of signal coverage. What would this plot look like if we moved the fixed location point on the other side of the building at (X=33, Y=3)?

```{r}
# Let's focus on the fixed point (33,3)
df = subset(offline, posX == 33 & posY == 3 & !(mac %in% remainingAP), c(signal, angle, mac))
plotBoxplotSignalStrength(df)
# figure 3 - Signal strength distribution for a point at the other end of the building
```
This plot is similar to the previous distribution but with no outliers. At this point we know signal strength is different for the two access points. We don't know if this will affect our prediction accuracy. Now, let us analyze the two AP's signal strength distribution over angles.

@@

```{r, fig.height=10}
df = subset(offline, posX == 24 & posY == 4 & !(mac %in% remainingAP), c(signal, angle, mac))
plotDensitySignalStrength(df)
# figure 4 - Signal strength distribution across angles
```
The above plot, shows the rejectedAP "dd:cd" has a higher signal density at lower strength while the acceptedAP "e1:c0" has a higher density coverage at a higher strength. The distributions show that many of the plots are normally distributed while some of them; i.e., "dd:cd" at 180 and "e1:c0" at 315 are heavily skewed or even bimodal. Looking at each AP from top to bottom, there seems to exist an obvious pattern with respect to angle. This would suggest the median of each distribution varies by angle which can be attributed to a difference in perceived strength. This means the angle of the device contributes to the APs strength of signal.

#### Average Signal Strength Distributions
Plot the signal strength distribution for all locations through the average signal strength for all locations by angle. Summary statistics for all possible locations by angle was created.

@@

```{r}
df = subset(offlineSummary, mac %in% c(chosenAP, rejectedAP), c(sdSignal, avgSignal, mac))
plotStdDevSignalStrength(df)
# figure 5 - Standard deviation of average signal strength
```

This above figure shows that there is considerable overlap between the two distributions of signal strength. This suggest that there may be value in keeping both APs in the k-NN model.

@@

Reshape the test data where each AP is in a column. 
```{r}
onlineSummary = castOnline(online)
#confirm rejected AP is included
dim(onlineSummary)
names(onlineSummary)
# Table 6 - Reshaped test data set
```

@@

<font color = "Red">talk about cross-validation</font>

@@

  # loop through neighbors
  for (k in 1:K) {
    estFold = predXY(newSignals = testFold[ , 6:12],
                     newAngles = testFold[ , 4],
                     trainFold, numAngles = 3, k = k)
    err[k] = err[k] + calcError(estFold, actFold)
  }
}
# plot k to sum of squared errors
plotSSErrors(err, K)
# figure 6 - estimated ideal 'k'
```
Given the plot above, the best k shows 9, but this changes if you rerun the cross-validation. Experimentation with k yielded 5 with the lowest error of 210 so it was used below.

@@

![Figure 7 - Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)

![Figure 8 - Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)

@@

![Figure 9 - Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)

@@

![*Figure 10 - Signal strength from AP @ 1,14*](SignalStrength1.png)

@@

![*Figure 11 - Signal strength from AP @ 33.5,2.8*](SignalStrength2.png)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Table of Contents: {#TOC}
<li>[Assignment Requirements](#AssignmentOutline)</li>
<li>[Introduction](#Introduction)</li>
<li>[**Part 1**](#Part1)</li>
<ul> 
  <li>[Signal Collection](#SigCol)</li>
  <li>[Signal Strength](#SigStren)</li>
  <li>[Signal Density](#SigDen)</li>
  <li>[Discussion](#Discussion)</li>
  <li>[k-Nearest Neighbors](#knn)</li>
</ul>
<li>[**Part 2**](#Part2)</li>
<ul> 
  <li>[2.1 Examining Power](#ExamPower)</li>
  <li>[2.2 Examining Weights](#ExamWeight)</li>
  <li>[2.3 Weighted kNN](#WtKNN)</li>
<li>[Future Study](#Future)</li>
</ul>
<li>[References](#References)</li>
<li>[Appendix](#Appendix)</li>

@@

@@

This is an example of the training input:

t=1139643118358;id=00:02:2D:21:0F:33; pos=0.0,0.0,0.0; degree=0.0; 00:14:bf:b1:97:8a=-38,2437000000,3; 00:14:bf:b1:97:90=-56,2427000000,3; 00:0f:a3:39:e1:c0=-53,2462000000,3; 00:14:bf:b1:97:8d=-65,2442000000,3; 00:14:bf:b1:97:81=-65,2422000000,3; 00:14:bf:3b:c7:c6=-66,2432000000,3; 00:0f:a3:39:dd:cd=-75,2412000000,3; 00:0f:a3:39:e0:4b=-78,2462000000,3; 00:0f:a3:39:e2:10=-87,2437000000,3; 02:64:fb:68:52:e6=-88,2447000000,1; 02:00:42:55:31:00=-84,2457000000,1

|Data Element|Description|Type|Example|
|:-----:|:------------------------------------:|:-----------------------------:|:---------:|
| t | timestamp (milliseconds since midnight 1/1/1970) |long integer|t=1139643118358 --> 2/11/2006 07:35:28|
|id|target device MAC address|6-octet sequence in hexadecimal|id=00:02:2D:21:0F:33|
|pos|coordinates of the target device in meters from a choke point in the host building|3 floating point values, comma-separated|pos=0.0,0.0,0.0|
|degree|Orientation of the target device (with respect to North) in degrees|float|degree=0.0|
|MAC Address|one or more reports from various Access Points|comma-separated value list|00:14:bf:b1:97:8a=-38,2437000000,3|

|Data Element|Description|Type|Example|
|:--|:----------------------------------------:|:--|:--:|
|signal|The signal strength reported for the Access Point in dBm|integer|-38|
|channel|The frequency of the channel used by the Access Point in Hz|integer|2437000000|
|mode|The mode of operation for the Access Point (1 = Ad Hoc, 3 = Access Point)|integer|3|

Six APs were initially discovered in the original analysis in the case study. The authors then decided to reject AP "00:0f:a3:39:dd:cd" (rejectedAP) stating it was identified as an extra address. The details of their reasoning was not clearly explained. The authors chose AP "00:0f:a3:39:e1:c0" (chosenAP) instead. This case study will evalute the author's decision and decide if the IPS would be more accurate with the chosenAP, rejectedAP, or both APs.

@@

![Figure 1 - An example of k-Nearest Neighbors from scikit-learn [2](#REF02)](knnExample.png)

In part 1 of the case study, the author's original decision to reject an AP is revisited to see if this was the best decision. In part 2, the prediction of X,Y based on the average of it's neighbors is revisited and modified to a weighted sum of X,Y to determine if this change will produce better predictions. Alternate weights will also be analyzed.

@@

```{r}
plotSignalMaps(c(chosenAP, rejectedAP))
#  figure 2 - number of reports at each location for the rejected and chosen APs.
```

@@

```{r}
df = subset(offline, posX ==2 & posY == 12 & !(mac %in% remainingAP), c(signal, angle, mac))
plotBoxplotSignalStrength(df)
# figure 3 - Signal strength distribution across both Access Points
```

The higher the decibel-milliwatts (dBm) means a stronger signal strength which indicates a closer physical position to the AP since signal strength decreases exponentially with distance. The distribution shows AP "e1:c0" has a stronger signal strength at this location than "dd:cd." But, the two of them together cover a wider range of signal coverage. What would this plot look like if we moved the fixed location point on the other side of the building at (X=33, Y=3)?

```{r}
# Let's focus on the fixed point (33,3)
df = subset(offline, posX == 33 & posY == 3 & !(mac %in% remainingAP), c(signal, angle, mac))
plotBoxplotSignalStrength(df)
# figure 4 - Signal strength distribution for a point at the other end of the building
```
This plot is similar to the previous distribution but with no outliers. At this point we know signal strength is different for the two access points. We don't know if this will affect our prediction accuracy. Now, let us analyze the two AP's signal strength distribution over angles.

@@

```{r, fig.height=10}
df = subset(offline, posX == 24 & posY == 4 & !(mac %in% remainingAP), c(signal, angle, mac))
plotDensitySignalStrength(df)
# figure 5 - Signal strength distribution across angles
```
The above plot shows the rejectedAP "dd:cd" has a higher signal density at lower strength while the acceptedAP "e1:c0" has a higher density coverage at a higher strength. The distributions show many of the plots are normally distributed while some of them, (e.g., "dd:cd" at 180 and "e1:c0" at 315) are heavily skewed or even bimodal. Looking at each AP from top to bottom, there seems to exist an obvious pattern with respect to angle. This would suggest the median of each distribution varies by angle which can be attributed to a difference in perceived strength. This means the angle of the device contributes to the APs strength of signal.

#### Average Signal Strength Distributions
Plot the signal strength distribution for all locations through the average signal strength for all locations by angle. Summary statistics for all possible locations by angle were created.

@@

```{r}
df = subset(offlineSummary, mac %in% c(chosenAP, rejectedAP), c(sdSignal, avgSignal, mac))
plotStdDevSignalStrength(df)
# figure 6 - Standard deviation of average signal strength
```

This above figure shows there is considerable overlap between the two distributions of signal strength. This suggest there may be value in keeping both APs in the k-NN model.

@@

Let's Reshape the test data where each AP is in a column. 
```{r}
onlineSummary = castOnline(online)
#confirm rejected AP is included
dim(onlineSummary)
names(onlineSummary)
# Table 6 - Reshaped test data set
```

@@

@@

  # loop through neighbors
  for (k in 1:K) {
    estFold = predXY(newSignals = testFold[ , 6:12],
                     newAngles = testFold[ , 4],
                     trainFold, numAngles = 3, k = k)
    err[k] = err[k] + calcError(estFold, actFold)
  }
}
# plot k to sum of squared errors
plotSSErrors(err, K)
# figure 7 - estimated ideal 'k'
```
Given the plot above, the best k shows 9, but this changes if you rerun the cross-validation. Experimentation with k yielded 5 with the lowest error of 210. Thus, it was used below.

@@

![Figure 8 - Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)

![Figure 9 - Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)

@@

![Figure 10 - Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)

@@

![*Figure 11 - Signal strength from AP @ 1,14*](SignalStrength1.png)

@@

![*Figure 12 - Signal strength from AP @ 33.5,2.8*](SignalStrength2.png)

@@

## <a id="Future">Future Study</a>
To take our work further, something to consider for a future endeavour would be to analyze the results of the K-Nearest Neighbors analysis for each Chosen and Rejected Access Point by the use of an F-test. The results from this study would help to support our findings in this case study by accepting or rejecting the null hypothesis ($H_O$) of no statistical difference between the results. Once these results are discovered, an analysis of variance could be conducted as well to put the nail in the coffin.
